---
title: "Comparing PhenoPath to treating covariates separately"
author: "Kieran Campbell"
date: "September 2017"
output:
  html_document:
    toc: yes
---

# Introduction

In this vignette we briefly compare PhenoPath to pseudotime inference on the Shalek et al. dataset splitting the covariates separately.



First load some useful libraries:

```{r, warning = FALSE}
suppressPackageStartupMessages(library(scater))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(monocle))
suppressPackageStartupMessages(library(phenopath))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(magrittr))

knitr::opts_chunk$set(fig.height = 4, fig.width = 6, fig.align = 'center',
                      cache = TRUE)

theme_set(theme_cowplot(font_size = 11))

set.seed(5L)
```

Function to fit pseudotimes with Monocle 2:

```{r}
fit_monocle2 <- function(sce) {
  exprs_mat <- exprs(sce)
  cds <- newCellDataSet(exprs_mat)
  sizeFactors(cds) <- rep(1, ncol(cds))
  cds <- setOrderingFilter(cds, rownames(exprs_mat))
  cds <- reduceDimension(cds, norm_method = "none")
  cds <- orderCells(cds)
  cds$time <- sce$time
  cds
}
```

Function to get $R^2$ compared to true time:

```{r}
get_R2 <- function(pseudotime, time) {
  time <- as.numeric(gsub("h", "", time))
  fit <- lm(pseudotime ~ time)
  s <- summary(fit)
  s$r.squared
}
```

Load the SCESet:

```{r}
sce <- readRDS("../../data/paper-scesets/sce_shalek_qc.rds")
```

Get highly variable genes and split by stimulant status:

```{r}
row_vars <- matrixStats::rowVars(exprs(sce))
high_var <- row_vars >= sort(row_vars, decreasing = TRUE)[5000]

sce_hvg <- sce[high_var, ]

sce_lps <- sce_hvg[, sce$stimulant == "LPS"]
sce_pam <- sce_hvg[, sce$stimulant == "PAM"]

sces <- list(
  lps = sce_lps,
  pam = sce_pam
)
```

Now fit the Monocle pseudotimes:

```{r}
cds_sets <- lapply(sces, fit_monocle2)
pseudotimes <- lapply(cds_sets, function(cds) cds$Pseudotime)
```

```{r}
print(get_R2(pseudotimes[[1]], sces[[1]]$time))
print(get_R2(pseudotimes[[2]], sces[[2]]$time))
```

And do the same for PhenoPath:

```{r}
pp_fit <- phenopath(sce_hvg, sce_hvg$x)

get_R2(trajectory(pp_fit), sce_hvg$time)
```

Let's go through the Monocle fits and perform differential expression:

```{r}
de_cds <- function(cds) {
  apply(exprs(cds), 1, function(y) {
    fit <- lm(y ~ cds$Pseudotime)
    s <- summary(fit)
    p_value <- coef(s)[2, c("Estimate", "Pr(>|t|)")]
    t(p_value)
  })
}

des <- lapply(cds_sets, de_cds)

beta_diff <- des[[1]][1,] - des[[2]][1,]
pvals <- lapply(des, function(d) d[2,])
qvals <- lapply(pvals, p.adjust, method = "BH")
log10qvals <- lapply(qvals, function(q) -log10(q))

qval_diff <- log10qvals[[1]] - log10qvals[[2]]
```

How does this compare to our PhenoPath fit?

```{r}
initial_plot <- qplot(beta_diff, pp_fit$m_beta[1,]) +
  labs(x = expression(beta[LPS]-beta[PAM]~"from linear model"),
       y = expression("PhenoPath"~beta))
print( initial_plot )
```

We notice it's completely off - because one of the pseudotimes is orientated the wrong way. Let's reverse this and repeat

```{r}
for(i in 1:2) {
  cds <- cds_sets[[i]]
  time_numeric <- as.numeric(gsub("h", "", cds$time))
  if(cor(cds$Pseudotime, time_numeric) < 0) {
    cds_sets[[i]]$Pseudotime <- max(cds$Pseudotime) - cds$Pseudotime
  }
}
```

Re-do DE analysis:

```{r}
des2 <- lapply(cds_sets, de_cds)

beta_diff2 <- des2[[1]][1,] - des2[[2]][1,]
pvals2 <- lapply(des2, function(d) d[2,])
qvals2 <- lapply(pvals2, p.adjust, method = "BH")
log10qvals2 <- lapply(qvals2, function(q) -log10(q))

qval_diff2 <- log10qvals2[[1]] - log10qvals2[[2]]
```

```{r}
reversed_plot <- qplot(beta_diff2, pp_fit$m_beta[1,]) +
  labs(x = expression(beta[LPS]-beta[PAM]~"from linear model"),
       y = expression("PhenoPath"~beta))
print( reversed_plot )
```

Save as supplementary figure:

```{r}
plot_grid(initial_plot, reversed_plot,
          nrow = 1, labels = 'AUTO')

ggsave("../../figs/supp_split_cov.png", width = 7, height = 3)
```


