---
title: "PhenoPath with arbitrary design matrices"
author: "Kieran Campbell"
date: "September 2017"
output:
  html_document:
    toc: yes
---

# Introduction

In this vignette we briefly introduce how to use PhenoPath with more complex design matrices. Rather than a binary or continuous variable, we consider a 3-level categorical variable.

# Simulating data

First load some useful libraries:

```{r, warning = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(phenopath))
suppressPackageStartupMessages(library(cowplot))

theme_set(theme_cowplot(font_size = 11))

knitr::opts_chunk$set(fig.height = 4, fig.width = 6, fig.align = 'center')

set.seed(5L)
```

We'll simulate for 200 samples and 80 features:

```{r}
N <- 200
G <- 80
```

First, draw `N` pseudotimes from a standard normal distribution:

```{r}
pst <- rnorm(N)
```

We'll sample the `x` variable from one of 3 levels, and turn that into a one-hot encoded design matrix using the `model.matrix` function:

```{r}
x <- factor(sample(1:3, N, replace = TRUE))
x_mat <- as.matrix(model.matrix(~ 0 + x))
head(x_mat)
```

Note that we use the no-intercept version of the design matrix (rather than `model.matrix(~ x)`) to make interpreting coefficients easier.

Next we'll create the $\alpha$ and $\beta$ matrix. For the purpose of illustration, we construct $\beta$ in a slightly unusual way so that each gene is affected only when $x$ takes on one of its three different levels for ease of interpretation:

```{r}
alpha <- matrix(rnorm(G * 3), ncol = 3)

lambda <- rnorm(G)

b <- sample(1:3, G, replace = TRUE)
beta <- as.matrix(model.matrix(~ 0 + factor(b)))
beta_sign <- matrix(sample(c(-1, 1), G * 3, replace = TRUE), ncol = 3)
beta <- beta * beta_sign
```

Finally we construct the expression matrix itself and add N(0,1) noise.

```{r}
Y <- x_mat %*% t(alpha) + x_mat %*% t(beta) * pst + pst %*% t(lambda)
Y <- apply(Y, 2, function(y) rnorm(N, y, 1))
```

We can quickly view this data in PCA space to get a feel for what it looks like:

```{r}
pca_df <- prcomp(Y)$x[,1:2] %>% 
  as_data_frame() %>% 
  mutate(pst, x)

by_pseudotime <- ggplot(pca_df, aes(x = PC1, y = PC2, color = pst)) +
  geom_point() +
  scale_colour_viridis(name = "Pseudotime")

by_x <- ggplot(pca_df, aes(x = PC1, y = PC2, color = x)) +
  geom_point() +
  scale_colour_brewer(palette = "Set1")

plot_grid(by_pseudotime, by_x, labels = "AUTO", nrow = 1)

ggsave("../../figs/supp_arbitrary_pca.png", width = 7, height = 3)
```

# Inference with PhenoPath

To perform inference with PhenoPath we simply pass in the matrix of gene expression and the one-hot-encoded design matrix. Although this can be done in other ways this leads to the easiest interpretation in terms of parameter estimates and posterior intervals:

```{r, cache = TRUE}
m <- as.matrix(model.matrix(~ 0 + factor(x)))
fit <- phenopath(Y, m)
```

And plot the ELBO to make sure things have converged:

```{r}
plot_elbo(fit)
```


# Inspecting results

First let's compare the inferred pseudotime against the true:

```{r}
qplot(pst, trajectory(fit)) +
  labs(x = "True pseudotime", y = "PhenoPath z")

true_vs_z <- last_plot()
```

Next let's get the matrix of fitted coefficients:

```{r}
beta_fit <- t(fit$m_beta)
head(beta_fit)
```

We can compare these to the true values:

```{r, fig.height = 3}
sig_ints <- significant_interactions(fit, 3)

dfs <- lapply(1:3, function(i) {
  data_frame(true = beta[,i], inferred = beta_fit[,i], x = as.character(i),
             significant = sig_ints[,i])
})
df <- bind_rows(dfs)

ggplot(df, aes(x = true, y = inferred, color = significant)) +
  geom_point() +
  facet_wrap(~ x) +
  scale_colour_brewer(name = "PhenoPath\nsignificant", palette = "Paired") +
  labs(x = expression(beta~"true"), y = expression(beta~"PhenoPath estimate")) +
  theme(legend.position = "top")

beta_plot <- last_plot()
```

```{r}
plot_grid(true_vs_z, beta_plot, ncol = 1, labels = "AUTO")
ggsave("../../figs/supp_arbitrary_results.png", width = 5, height = 6)
```


Finally, we can look at the significant interactions:

```{r, fig.width = 4}
misclass <- as_data_frame(abs(beta) - 1 * significant_interactions(fit, 3))
names(misclass) <- as.character(1:3)
misclass$g <- seq_len(nrow(misclass))

misclass_tidy <- gather(misclass, p, classif, -g)
misclass_tidy$classif <- plyr::mapvalues(misclass_tidy$classif,
                                         from = c(-1, 0, 1),
                                         to = c("False positive", "Correct", "False negative"))

ggplot(misclass_tidy, aes(x = p, y = g, fill = classif)) +
  geom_raster() +
  scale_fill_brewer(palette = "Blues")
```


# Technical

```{r}
sessionInfo()
```



